---
title: "Assignment-7"
author: "Thomas McHale"
date: "06 April 2023"
echo: true
warning: false
format: html
editor: visual
---

# Assignment 7

### Date and Time: `r Sys.time()`

```{r}
source("sepsis_monitor_functions.R")

library(dplyr)
library(tidyr)
library(knitr)
library(ggplot2)
library(tidyverse)
library(data.table) ## For the fread function
library(lubridate)
library(tictoc)
library(googledrive)

drive_deauth()
file_link <- "https://drive.google.com/file/d/1-BqFL73u2z4Uq-bGZbmIfrpATnNr4ymu/view?usp=share_link"

## All data up until now
new_data <- updateData(file_link)

## Include only most recent data
most_recent_data <- new_data %>%
  group_by(PatientID) %>%
  filter(obsTime == max(obsTime))

```

The patients we will look at in this report include the last 50 observations from the time the report was started, indicated above. This includes `r n_distinct(most_recent_data$PatientID)` patients recently in the ICU. Of those `r n_distinct(most_recent_data$PatientID)` patients, `r n_distinct(most_recent_data$SepsisLabel==1)` have sepsis at the time of this report. We will start by looking at the vitals of patients who currently have sepsis.

### Table 1: Most recent vital signs of patients in the ICU with sepsis

```{r}
sepsis_pts <- subset(most_recent_data, SepsisLabel==1)

sepsis_table <- select(sepsis_pts, -ICULOS, -SepsisLabel, -obsTime)
colnames(sepsis_table) <- c("Patient ID", "Heart Rate", "Temperature", "Respiratory Rate")
kable(sepsis_table)

```

### Figure 1: Vitals signs for patients with sepsis while in the ICU

```{r}

sepsis_plots <- lapply(unique(sepsis_pts$PatientID), function(patient){
  patient_data <- subset(sepsis_pts, PatientID == patient)
  ggplot(data = patient_data, aes(x = obsTime)) +
    geom_point(data = subset(patient_data, !is.na(HR)), aes(y = HR, colour = "HR")) +
    geom_point(data = subset(patient_data, !is.na(Temp)), aes(y = Temp, colour = "Temp")) +
    geom_point(data = subset(patient_data, !is.na(Resp)), aes(y = Resp, colour = "Resp")) +
    labs(x = "Observation Time", y = "Resp = breaths/minute,  Temp = ºC,  HR = beats/minute", colour = "Vital Sign") +
    scale_x_datetime(date_labels = "%m/%d/%Y %H:%M") + 
    ggtitle(paste("Patient", patient))
})
sepsis_plots

```

### Table 2: Recent Change in Vital Signs

```{r}
sepsis_pts <- subset(new_data, SepsisLabel==1)

sepsis_pts %>%
  group_by(PatientID) %>%
  arrange(obsTime) %>%
  mutate(prevHR = lag(HR), prevTemp = lag(Temp), prevResp = lag(Resp)) %>%
  summarise(
    `Change in Heart Rate (bpm)` = last(HR) - last(prevHR),
    `Change in Temp (ºC)` = last(Temp) - last(prevTemp),
    `Change in Resp Rate (rpm)` = last(Resp) - last(prevResp)
  ) %>%
  kable()
```

::: callout-note
Note that the most_recent_data file only had a single observation of vitals for patients. Thus in order for the change in vitals to display meaningful data, I switched to the new_data file.

The plots are also more interesting when looking at new_data vs most_recent_data (and changing to a line instead of point). But I leave it as most_recent_data. The code for this is included in my script.
:::
